---
title: "R Notebook for SDM of catch species in the Mexican Pacific and Gulf of Mexico"
output: html_notebook
---

 

```{r Attach packages, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# List of packages for session
source("libraries.R")

source("fisheries_sdm_CONAPESCA_chunks.R")
source("update_species.R")
source("thin_data.R")
source("make_occ_map.R")
source("make_all_map.R")
source("ensemble_smd.R")
source("ensemble_project.R")
source("add_anomaly.R")
```

```{r}
#create map
#scale factor is how much distance to add in the edges of the map, units are dependent on the coordinate system of the shapefile
#bar position can be top or bottom (t,b) or left or right (l,r)

model.map <- make_map(shape.file="ne_10m_land.shp", file.name = "mexico_ezz_map.png", scale.factor = 2, bar.position = "tl",
                      min.long= -130.00, max.long=-50, min.lat=-10.00, max.lat=40.00)

model.map
```

Manually reviewed species lists, to fix outstanding synonyms or taxonomic changes. Then updated the coordinate files
```{r}

update_species(thissamplearea="atlantic")

update_species(thissamplearea="pacific")



```


We thinned the occurrence records in geographic space to correct for this bias; this approach takes a set of occurrence records and identifies multiple new subsets that meet the minimum nearest neighbor distance constraint, based on an automated randomization approach


```{r thindata, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
datapath <- "~/conapescaspecies" #put path
setwd(datapath)

# thissamplearea <- "atlantic"
# 
# goc.20.sp <- read_csv(paste(thissamplearea,"_CONAPESCA_sp_20reg_rev.csv"))
# 
# sp.goc.20 <- goc.20.sp %>% 
#     distinct(species) %>% 
#     arrange(species) %>% 
#     pull(species)
# 
# 
# mclapply(sp.goc.20,thin_data, goc.20.sp, thissamplearea)

thissamplearea  <- "pacific"

goc.20.sp <- read_csv(paste(thissamplearea,"_CONAPESCA_sp_20reg.csv"))

sp.goc.20 <- goc.20.sp %>% 
    distinct(species) %>% 
    arrange(species) %>% 
    pull(species)

 
mclapply(sp.goc.20,thin_data, goc.20.sp, thissamplearea)

```


We only had max, min, and mean for observed rasters, so calculated the range
```{r}
lapply(sample.areas,get_range)
setwd(datapath)

```

#retrieved rasters from Bio-Oracle
```{r getbiooracle, message=FALSE, warning=FALSE, include=FALSE}
crs.geo.wgs <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")  # geographical, datum WGS84

biooracle.rasters <- list_layers(c("Bio-ORACLE")) %>% 
  as_tibble()

biooracle.rasters %>% 
  write_csv("biooracle_layers.csv")

biooracle.codes <- biooracle.rasters %>% 
  filter(!grepl("ice",layer_code)) %>% 
   filter(!grepl("cloud",layer_code)) %>% 
  distinct(layer_code)

# (down)load specific layers 
biooracle.data <- load_layers(biooracle.codes)

list.layers <- list.files("~/conapescaspecies/biooracle", pattern="*.*tif$", full.names = TRUE)

save_dir <- "~/conapescaspecies/biooracle_asc"
dir.create(save_dir)

crs.geo.wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"  # geographical, datum WGS84

reference.raster <- raster("~/conapescaspecies/rasters/atlantic/alk_max.asc")
projection(reference.raster) <-CRS(crs.geo.wgs)  # geographical, datum WGS84
 
#cut rasters to study area  
lapply(list.layers,cut_raster,reference.raster,crs.geo.wgs, study.area="atlantic")

reference.raster <- raster("~/conapescaspecies/rasters/pacific/alk_max.asc")
projection(reference.raster) <-CRS(crs.geo.wgs)  # geographical, datum WGS84
 
#cut rasters to study area  
lapply(list.layers,cut_raster,reference.raster,crs.geo.wgs, study.area="pacific")

```

Estimated colinearity of observed rasters, there were too many layers, so this helped narrow down the dataset

```{r colinear_test, message=FALSE, warning=FALSE, include=FALSE}

#load first set of rasters

thissamplearea = "pacific"

list.rasters <- list.files(path="~/conapescaspecies/biooracle_asc/",pattern = paste0("*.*",thissamplearea,".asc"), full.names = TRUE)

colinear_test(thissamplearea, parameter.names = "biooracle_layers.csv", list.rasters, raster.name="biooracle")


thissamplearea = "atlantic"

list.rasters <- list.files(path="~/conapescaspecies/biooracle_asc/",pattern = paste0("*.*",thissamplearea,".asc"), full.names = TRUE)

colinear_test(thissamplearea, parameter.names = "biooracle_layers.csv", list.rasters, raster.name="biooracle")


thissamplearea = "pacific"

list.rasters <- list.files(path=paste("~/conapescaspecies/rasters/",thissamplearea,sep=""), pattern=".asc", full.names = TRUE)

colinear_test(thissamplearea, parameter.names = "observed_layers.csv", list.rasters, raster.name="observed")


thissamplearea = "atlantic"

list.rasters <- list.files(path=paste("~/conapescaspecies/rasters/",thissamplearea,sep=""), pattern=".asc", full.names = TRUE)

colinear_test(thissamplearea, parameter.names = "observed_layers.csv", list.rasters, raster.name="observed")


  
```


Created projected rasters
```{r}

#MPI model
obs.dir <- "~/conapescaspecies/rasters/pacific"
save.dir <- "~/conapescaspecies/rasters/pacific_mpi"

thismodel = "MPI_ESM"

cmip.rasters <- list.files(path="~/climatemodels/climatemodelensembles/",pattern = "tos*.*MPI-ESM-MR*.*rcp85*.*2070", full.names = TRUE)

all.obs.rasters <- list.files(path=obs.dir,pattern ="sst*.*")
                               
observed.rasters <- all.obs.rasters[!grepl("range",all.obs.rasters)]

lapply(observed.rasters,add_anomaly,cmip.rasters, save.dir, obs.dir, thismodel)


cmip.rasters <- list.files(path="~/climatemodels/climatemodelensembles/",pattern = "sos_*.*MPI-ESM-MR*.*rcp85*.*2070", full.names = TRUE)

all.obs.rasters <- list.files(path=obs.dir,pattern ="sal*.*")
                               
observed.rasters <- all.obs.rasters[!grepl("range",all.obs.rasters)]

lapply(observed.rasters,add_anomaly,cmip.rasters, save.dir, obs.dir, thismodel)


#GFDL model

obs.dir <- "~/conapescaspecies/rasters/pacific"
save.dir <- "~/conapescaspecies/rasters/pacific_gfdl"

thismodel = "GFDL"

list.rasters <- list.files(path="~/climatemodels/climatemodelensembles/",pattern = "tos*.*GFDL-ESM2M_*.*rcp85*.*2070", full.names = TRUE)


all.obs.rasters <- list.files(path=obs.dir,pattern ="sst*.*")
                               
observed.rasters <- all.obs.rasters[!grepl("range",all.obs.rasters)]

lapply(observed.rasters,add_anomaly,cmip.rasters, save.dir, obs.dir, thismodel)


list.rasters <- list.files(path="~/climatemodels/climatemodelensembles/",pattern = "sos*.*GFDL-ESM2M_*.*rcp85*.*2070", full.names = TRUE)


all.obs.rasters <- list.files(path=obs.dir,pattern ="sal*.*")
                               
observed.rasters <- all.obs.rasters[!grepl("range",all.obs.rasters)]

lapply(observed.rasters,add_anomaly,cmip.rasters, save.dir, obs.dir, thismodel)


```


Plot occurrence
```{r}

map.dir <- "/home/atlantis/conapescaspecies/occurrence_maps/"
  dir.create(map.dir)
  
thissamplearea <- "pacific"

sp.list <- list.files(path=paste("~/conapescaspecies/thindata",thissamplearea,sep="_"), pattern = "*.*txt$") %>% 
  gsub("_thinned_full_log_file.txt","",.)

pacific.rec <- lapply(sp.list,make_occ_map,thissamplearea,min.lon= -130.00, max.lon=-50, min.lat=-10.00, max.lat=40.00, map.dir)

pacific.occ <- pacific.rec %>% 
  discard(is.null) %>% 
  bind_rows()

write_csv(pacific.occ,"pacific_occ_thinned.csv")


thissamplearea <- "atlantic"

sp.list <- list.files(path=paste("~/conapescaspecies/thindata",thissamplearea,sep="_"), pattern = "*.*txt$") %>% 
  gsub("_thinned_full_log_file.txt","",.)

atlantic.rec <- lapply(sp.list,make_occ_map,thissamplearea,min.lon= -130.00, max.lon=-50, min.lat=-10.00, max.lat=40.00, map.dir)

atlantic.occ <- atlantic.rec %>% 
  discard(is.null) %>% 
  bind_rows()

write_csv(atlantic.occ,"atlantic_occ_thinned.csv")


combined.data <- atlantic.occ %>% 
  bind_rows(pacific.occ)
  
make_all_map(combined.data, min.lon= -130.00, max.lon=-50, min.lat=-10.00, max.lat=40.00)


```


Plot species composition

```{r}
sp.rec <- read_csv("pacific_occ_thinned.csv") %>% 
  group_by(species) %>% 
  summarise(no_records = n()) %>% 
 # filter(no_records > 100) %>% 
  arrange(desc(no_records)) %>% 
  mutate(sum_abundance = cumsum(no_records), tot_rec = sum(no_records), prop_records = (sum_abundance/tot_rec)*100) %>% 
  filter(!species%in%c("oxyporhamphus micropterus micropterus","oxyporhamphus micropterus")) %>% 
  filter(prop_records<20) %>% 
  mutate(species = str_to_sentence(species))

ggplot() +
       geom_col(data = sp.rec , aes(x = species, y = prop_records), fill="darkblue") +
       scale_fill_viridis_c(na.value = "gray85") +
       coord_quickmap()+
       theme_minimal()+
       xlab("Species") + 
       ylab("Percent cumulative abundance") +
      theme(axis.text.x = element_text(angle = 90)) 
  
  ggsave("cumulative_abundance_plot.png", dpi=300)
      
```



Estimate ensemble models, observed rasters 

```{r get_esmd, message=TRUE, warning=TRUE, include=FALSE}

thissamplearea <- "pacific"

sp.list <- read_csv("pacific_occ_thinned.csv") %>% 
  group_by(species) %>% 
  summarise(no_records = n()) %>% 
  filter(no_records > 20) %>% 
  arrange(no_records) %>% 
  .$species


#sp.list <- list.files(path=paste("~/conapescaspecies/thindata",thissamplearea,sep="_"), pattern = "*.*txt$") %>% 
#  gsub("_thinned_full_log_file.txt","",.)


NumberOfCluster <- detectCores() - 4
cl <- makeCluster(NumberOfCluster)
registerDoSNOW(cl)
 
clusterExport(cl, c("ensemble_smd"))

raster.path ="~/conapescaspecies/rasters/pacific/"

thissamplearea <- "pacific"

#raster.name = "observed"
raster.name = "biooracle"

sdm.dir <- paste0("/home/atlantis/conapescaspecies/sdm_",thissamplearea,"_",raster.name)
dir.create(sdm.dir)

raster.sel.list <- read_csv(file=paste(thissamplearea,raster.name,"rasters_selected.csv",sep="_")) %>% 
      pull(x) %>% 
      paste(.,".asc",sep="") %>% 
      paste(raster.path,., sep="")
    
rasters.selected <- stack(raster.sel.list)
    
projection(rasters.selected) <-CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")  # geographical, datum WGS84
    
ocean.shp <- readOGR("ne_10m_ocean.shp")

 sp.esdm <- foreach(this.species=sp.list, .verbose = T, .export = c("raster.path","thissamplearea","raster.name","sdm.dir","ocean.shp")) %dopar% {


 .packages <- (c("tidyverse","raster","SSDM","cowplot"))

 # Install CRAN packages (if not already installed)

 if(!require(SSDM)){
      devtools::install_github("sylvainschmitt/SSDM")

 }

 .inst <- .packages %in% installed.packages()
 if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst], repos='http://cran.us.r-project.org', dependencies = TRUE)

 # Load packages into session
 lapply(.packages, require, character.only=TRUE)


#testing sp.list[1:5]

ensemble_smd(this.species,raster.name,thissamplearea,raster.path,rasters.selected, sdm.dir, ocean.shp)


}

#}

stopCluster(cl)

system("az vm deallocate --name atlantisserver06 --no-wait --resource-group morzariacedogroup")


```

Estimate ensemble models, biooracle rasters

```{r get_esmd, message=TRUE, warning=TRUE, include=FALSE}

thissamplearea <- "pacific"

sp.list <- read_csv(paste0(thissamplearea,"_occ_thinned.csv")) %>% 
  group_by(species) %>% 
  summarise(no_records = n()) %>% 
  filter(no_records > 20) %>% 
  arrange(no_records) %>% 
  .$species


#sp.list <- list.files(path=paste("~/conapescaspecies/thindata",thissamplearea,sep="_"), pattern = "*.*txt$") %>% 
#  gsub("_thinned_full_log_file.txt","",.)


NumberOfCluster <- detectCores() - 4
cl <- makeCluster(NumberOfCluster)
registerDoSNOW(cl)
 
clusterExport(cl, c("ensemble_smd"))

raster.path ="~/conapescaspecies/biooracle_asc/"

thissamplearea <- "pacific"

raster.name = "biooracle"

sdm.dir <- paste0("/home/atlantis/conapescaspecies/sdm_",thissamplearea,"_",raster.name)
dir.create(sdm.dir)

raster.sel.list <- read_csv(file=paste(thissamplearea,raster.name,"rasters_selected.csv",sep="_")) %>% 
      pull(x) %>% 
      paste(.,".asc",sep="") %>% 
      paste(raster.path,., sep="")
    
rasters.selected <- stack(raster.sel.list)
    
projection(rasters.selected) <-CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")  # geographical, datum WGS84
    
ocean.shp <- readOGR("ne_10m_ocean.shp")

 sp.esdm <- foreach(this.species=sp.list, .verbose = T, .export = c("raster.path","thissamplearea","raster.name","sdm.dir","ocean.shp")) %dopar% {


 .packages <- (c("tidyverse","raster","SSDM","cowplot"))

 # Install CRAN packages (if not already installed)

 if(!require(SSDM)){
      devtools::install_github("sylvainschmitt/SSDM")

 }

 .inst <- .packages %in% installed.packages()
 if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst], repos='http://cran.us.r-project.org', dependencies = TRUE)

 # Load packages into session
 lapply(.packages, require, character.only=TRUE)


#testing sp.list[1:5]

ensemble_smd(this.species,raster.name,thissamplearea,raster.path,rasters.selected, sdm.dir, ocean.shp)


}

#}

stopCluster(cl)

system("az vm deallocate --name atlantisserver06 --no-wait --resource-group morzariacedogroup")


```


#calculate areas with higher model uncertainty
```{r}

sp.list <- list.files(path="~/conapescaspecies/sdm_pacific_observed/", pattern = "uncertainty.tif$", full.names = TRUE, recursive = TRUE) 

tot.raster <- raster(sp.list[1])


for(this.file in 2:length(sp.list)){

  raster.name <- sp.list[this.file]
  print(raster.name)
    
  this.raster <- raster(raster.name)
  
  tot.raster <- this.raster + tot.raster
  
}

norm.tot.raster <- normImage(tot.raster, norm=FALSE)

norm.tot.raster <- raster.transformation(tot.raster, trans = "norm", smin = 0, smax = 255)

raster.plot <- ggplot() +
  geom_raster(data = norm.tot.raster , aes(x = x, y = y, fill = layer)) + 
  scale_fill_viridis_c(na.value="gray66")+
  coord_quickmap()


  ggsave("uncertainty_summary_observed.png", raster.plot, device ="png", dpi=300)

```



#calculate areas with higher probability
```{r}

sp.list <- list.files(path="~/conapescaspecies/sdm_pacific_observed/", pattern = "Probability.tif$", full.names = TRUE, recursive = TRUE) 

tot.raster <- raster(sp.list[1])


for(this.file in 2:length(sp.list)){

  raster.name <- sp.list[this.file]
  print(raster.name)
    
  this.raster <- raster(raster.name)
  
  tot.raster <- this.raster + tot.raster
  
}

norm.tot.raster <- normImage(tot.raster, norm=FALSE)

norm.tot.raster <- raster.transformation(tot.raster, trans = "norm", smin = 0, smax = 255)

writeRaster(norm.tot.raster,file="combined_raster_probability", format="ascii")

raster.plot <- ggplot() +
  geom_raster(data = norm.tot.raster , aes(x = x, y = y, fill = layer)) + 
  scale_fill_viridis_c(na.value="gray66")+
  coord_quickmap()


  ggsave("probability_summary_observed.png", raster.plot, device ="png", dpi=300)

```


Predict using future probability layers GFDL

```{r}

sp.list <- read_csv("pacific_occ_thinned.csv") %>% 
  group_by(species) %>% 
  summarise(no_records = n()) %>% 
  filter(no_records > 100) %>% 
  arrange(no_records) %>% 
  .$species

#sp.list <- list.files(path=paste("~/conapescaspecies/thindata",thissamplearea,sep="_"), pattern = "*.*txt$") %>% 
#  gsub("_thinned_full_log_file.txt","",.)


NumberOfCluster <- detectCores() - 2
cl <- makeCluster(NumberOfCluster)
registerDoSNOW(cl)
 
clusterExport(cl, c("ensemble_smd"))

raster.path ="~/conapescaspecies/rasters/pacific_gfdl/"

thismodel = "GFDL"
thissamplearea <- "pacific"

raster.name = "observed"

sdm.dir <- paste0("/home/atlantis/conapescaspecies/sdm_",thissamplearea,"_",raster.name)

raster.sel.list <- read_csv(file=paste(thissamplearea,raster.name,"rasters_selected.csv",sep="_")) %>% 
      pull(x) %>% 
      gsub("[.]","-",.) %>% 
      paste(.,".asc",sep="") %>% 
      paste(raster.path,., sep="")
    
rasters.selected <- stack(raster.sel.list)
    
projection(rasters.selected) <-CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")  # geographical, datum WGS84
    
ocean.shp <- readOGR("ne_10m_ocean.shp")

 sp.esdm <- foreach(this.species=sp.list, .verbose = T, .export = c("raster.path","thissamplearea","raster.name","sdm.dir","ocean.shp","rasters.selected","thismodel")) %dopar% {


 .packages <- (c("tidyverse","raster","SSDM","cowplot"))

 # Install CRAN packages (if not already installed)

 if(!require(SSDM)){
      devtools::install_github("sylvainschmitt/SSDM")

 }

 .inst <- .packages %in% installed.packages()
 if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst], repos='http://cran.us.r-project.org', dependencies = TRUE)

 # Load packages into session
 lapply(.packages, require, character.only=TRUE)


#testing sp.list[1:5]

ensemble_project(this.species,raster.name,thissamplearea,raster.path,rasters.selected, sdm.dir, ocean.shp, thismodel)


}

#}

stopCluster(cl)

system("az vm deallocate --name atlantisserver06 --no-wait --resource-group morzariacedogroup")


```

Predict using future probability layers MPI

```{r}

sp.list <- read_csv("pacific_occ_thinned.csv") %>% 
  group_by(species) %>% 
  summarise(no_records = n()) %>% 
 # filter(no_records > 100) %>% 
  arrange(no_records) %>% 
  .$species

#sp.list <- list.files(path=paste("~/conapescaspecies/thindata",thissamplearea,sep="_"), pattern = "*.*txt$") %>% 
#  gsub("_thinned_full_log_file.txt","",.)


NumberOfCluster <- detectCores() - 1
cl <- makeCluster(NumberOfCluster)
registerDoSNOW(cl)
 
clusterExport(cl, c("ensemble_smd"))

raster.path ="~/conapescaspecies/rasters/pacific_mpi/"

thismodel = "MPI-ESM"
thissamplearea <- "pacific"

raster.name = "observed"

sdm.dir <- paste0("/home/atlantis/conapescaspecies/sdm_",thissamplearea,"_",raster.name)

raster.sel.list <- read_csv(file=paste(thissamplearea,raster.name,"rasters_selected.csv",sep="_")) %>% 
      pull(x) %>% 
      gsub("[.]","-",.) %>% 
      paste(.,".asc",sep="") %>% 
      paste(raster.path,., sep="")
    
rasters.selected <- stack(raster.sel.list)
    
projection(rasters.selected) <-CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")  # geographical, datum WGS84
    
ocean.shp <- readOGR("ne_10m_ocean.shp")

 sp.esdm <- foreach(this.species=sp.list, .verbose = T, .export = c("raster.path","thissamplearea","raster.name","sdm.dir","ocean.shp","rasters.selected","thismodel")) %dopar% {


 .packages <- (c("tidyverse","raster","SSDM","cowplot"))

 # Install CRAN packages (if not already installed)

 if(!require(SSDM)){
      devtools::install_github("sylvainschmitt/SSDM")

 }

 .inst <- .packages %in% installed.packages()
 if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst], repos='http://cran.us.r-project.org', dependencies = TRUE)

 # Load packages into session
 lapply(.packages, require, character.only=TRUE)


#testing sp.list[1:5]

ensemble_project(this.species,raster.name,thissamplearea,raster.path,rasters.selected, sdm.dir, ocean.shp,thismodel)


}

#}

stopCluster(cl)

system("az vm deallocate --name atlantisserver06 --no-wait --resource-group morzariacedogroup")


```



#calculate areas with higher model probability under future climate change
```{r}

sp.list <- list.files(path="~/conapescaspecies/sdm_pacific_observed/", pattern = "projection.asc$", full.names = TRUE, recursive = TRUE) 

tot.raster <- raster(sp.list[1])


for(this.file in 2:length(sp.list)){

  raster.name <- sp.list[this.file]
  print(raster.name)
    
  this.raster <- raster(raster.name)
  
  tot.raster <- this.raster + tot.raster
  
}

norm.tot.raster <- normImage(tot.raster, norm=FALSE)

norm.tot.raster <- raster.transformation(tot.raster, trans = "norm", smin = 0, smax = 255)

writeRaster(norm.tot.raster,file="combined_raster_projection", format="ascii")


raster.plot <- ggplot() +
  geom_raster(data = norm.tot.raster , aes(x = x, y = y, fill = layer)) + 
  scale_fill_viridis_c(na.value="gray66")+
  coord_quickmap()


  ggsave("projection_summary_observed.png", raster.plot, device ="png", dpi=300)

```



```{r}

proj.raster <- raster("combined_raster_projection.asc")


prob.raster <- raster("combined_raster_probability.asc")


diff.raster <- prob.raster - proj.raster

norm.tot.raster <- raster.transformation(diff.raster, trans = "norm", smin = 0, smax = 255)

raster.plot <- ggplot() +
  geom_raster(data = norm.tot.raster , aes(x = x, y = y, fill = layer)) + 
  scale_fill_viridis_c(na.value="gray66")+
  coord_quickmap()


  ggsave("difference_summary_observed.png", raster.plot, device ="png", dpi=300)

```




